Community-Conditioned Language Models

### 2020-11-02 paper/paper.tex
add community infreence confusion plots
address some comments

\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}

@@

Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Our dataset includes \emph{comments}\jp{Comments/Post terminology; make sure its consistens} from \num{46} subreddits, 
randomly selected from the set of all forums 
with at least \num{15000} comments per month for each month
in the three years between 2015 and 2017. 
We initially selected \num{50} subreddits, 
but excluded four of them from further analysis: 
two which were primarily non-English and two with particularly\jp{quantify?} short average comment lengths.
Each community corpus consist of \num{50000} randomly selected comments from the year 2015.

Comments were preprocessed as follows: 
We excluded the content of block quotes, code blocks, and tables.
We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3\jp{citation?}
and lower-cased all tokens.

@@

At the the core of the community-conditioned language model (CCLM) 
is a standard $n$\jp{with n=3}-layer neural sequence encoder
with token embedding (we experiment with LSTM and transformer encoders).
To condition the model, we additionally train a \emph{community embedding} parameter.
This parameter is concatenated with the hidden layer of the sequence encoder
for some layer $c \leq n$ and passed through a linear layer, 
which projects the resulting vector back down to the origial hidden size.
For $c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $c=0$, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

We used\jp{past tense is a bit awkward?} a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consisted of the most frequent tokens at the comment level,
meaning that we only counted each token once per comment.
The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
many times in a single comment.

We trained the models on a simple left-to-right\jp{I don't think that this applies for the transformer model} language modeling task 
with cross entropy loss over the vocabulary 
and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens.\jp{What happens with longer sequences? Are sequences split at sentence boundaries?}
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

We trained the models until the validation loss stopped decreasing for two epochs in a row.
\jp{How many epochs does that turn out to be? Is there a meaningful difference between lstm/}
The weights from the epoch with the smallest validation loss were used for testing.

\begin{table}
  \centering
  \input{floats/best_epoch.tex}
  \label{tab:best-epoch}
  \caption{Epoch with the lowest validation loss.}
\end{table}

@@

TODO: For space reasons, it might be better to put the community-level table in the appendix
and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.

We also note that the conditioned models perform differently across different communities---\
for some communities, the test perplexity of the conditioned models is a lot lower,\jp{Where can I see this?}
whereas for others it is the same or even slightly higher.
(TODO: talk about this in terms of mutual information instead?)
As an anecdotal observation, we note 
that communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics.\jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.

@@

As a qualitative assesment of the community embeddings, 
we identified several topic categories into which at least three
of the selected communities fall. 
We plotted the two-component PCA projection of the embeddings,
annotated with these topic categories (Figure~\ref{comm-pca}).
We see that there is some evidence of clustering according to
the topic categories we identified,
suggesting that the embedding space learned by the model is meaningful.

@@

We now turn our interest to language-model-based community classifiers
(LMCC). Given a post $m$, and LMCC gives the probability that it
belongs in a community $c_i$ Using the Bayes theorem, we can turn a
CCLM into a LMCC. Indeed, the probability that a given post $m$
belongs to a community $c_i$ can be calculated as follows:
\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m | c_i)$ is given by the CCLM. $P(m)$ is our normalisation factor, which can be set so that
\[\sum_i P(c=c_i | m) = 1\].

Knowing that $m$ \emph{actually} comes from community $c_j$, we can
measure the the precision our classifier, as a confusion matrix $C$,
which can be calculated as follows for a given set of posts:(TODO: what set exactly?)
\[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].

@@

@@

While the confusion matrices provide a lot of insight, they are
somewhat \emph{too} fine-grained. For this reason, we will be
interested in confusing the posts from community $c_j$ are to the
LMCC. For this purpose, we consider the perplexity of the LMCC for a
post coming from community $j$, defined as the exponential of the
entropy of the distribution \(D_j\), which is the (discrete)
probability distribution of predicted community for a random post from
$c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
\[Ppl_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

@@

We see a positive correlation: this suggests that the more stable a
community is, the more difficult it is to identify posts as coming
from it. What this suggests is that stable communities tend to use a
varied and standard, subset of English, while unstable communities
resort to more formulaic posts. This may suggests that to 'fit' in an
unstable community, one must use more obvious linguistic cues than in
a stable community. In fact, it is reasonable to think that a new
member will ostensibly use community-specifc language, but once
established in a community, a member would tend not to do so.

>>>>

\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
\newcommand\bn[1]{\todo[backgroundcolor=green!10]{BN: #1}}

@@

Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.

Our dataset includes comments from \num{46} subreddits, 
randomly selected from the set of all forums 
with at least \num{15000} comments per month for each month
in the three years between 2015 and 2017. 
We initially selected \num{50} subreddits, 
but excluded four of them from further analysis: 
two which were primarily non-English and two with particularly\jp{quantify?} short average comment lengths.
Each community corpus consist of \num{50000} randomly selected comments from the year 2015.

Comments were preprocessed as follows: 
We excluded the content of block quotes, code blocks, and tables.
We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
and lower-cased all tokens.

@@

At the the core of the community-conditioned language model (CCLM) 
is a standard $n$-layer neural sequence encoder
with token embedding (we experiment with LSTM and transformer encoders and set $n=3$ for all models).
To condition the model, we additionally train a \emph{community embedding} parameter.
This parameter is concatenated with the hidden layer of the sequence encoder
for some layer $c \leq n$ and passed through a linear layer, 
which projects the resulting vector back down to the origial hidden size.
For $c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $c=0$, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consisted of the most frequent tokens at the comment level,
meaning that we only counted each token once per comment.
The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
many times in a single comment.

We trained the models on a simple left-to-right\jp{I don't think that this applies for the transformer model} language modeling task 
with cross entropy loss over the vocabulary 
and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer comments (16.8\% of comments were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

We trained the models until the validation loss stopped decreasing 
for two epochs in a row,
and used the weights from the epoch with the smallest validation loss for testing.
In general, transformer models trained for about half as long as the LSTM models (Table~\ref{tab:best-epoch}).

\begin{table}
  \centering
  \input{floats/best_epoch.tex}
  \caption{Epoch with the lowest validation loss.}
  \label{tab:best-epoch}
\end{table}

@@

%TODO: For space reasons, it might be better to put the community-level table in the appendix and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.}

We also note that the conditioned models perform differently across different communities---\
for some communities, the test perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
whereas for others it is the same or even slightly higher.
\bn{TODO: talk about this in terms of mutual information instead?}
As an anecdotal observation, we note 
that communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics.\jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.

@@

As a qualitative assesment of the community embeddings, 
we identified several topic categories into which at least three
of the selected communities fall. 
We plotted the two-component PCA projection of the embeddings,
annotated with these topic categories (Figure~\ref{fig:comm-pca}).
We see that there is some evidence of clustering according to
the topic categories we identified,
suggesting that the embedding space learned by the model is meaningful.

@@

We now turn our interest to language-model-based community classifiers
(LMCC). Given a comment $m$, and LMCC gives the probability that it
belongs in a community $c_i$ Using the Bayes theorem, we can turn a
CCLM into a LMCC. Indeed, the probability that a given comment $m$
belongs to a community $c_i$ can be calculated as follows:
\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
In the above,
$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
$P(m | c_i)$ is given by the CCLM. $P(m)$ is our normalisation factor, which can be set so that
\[\sum_i P(c=c_i | m) = 1\].

Knowing that $m$ \emph{actually} comes from community $c_j$, we can
measure the precision our classifier, as a confusion matrix $C$,
which can be calculated as follows for a given set of comment (TODO: what set exactly?)
\[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].

@@

\begin{figure*}
\begin{tikzpicture}
  \confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
\end{tikzpicture}
\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:confusion}
\end{figure*}

@@

While the confusion matrices provide a lot of insight, they are
somewhat \emph{too} fine-grained. For this reason, we will be
interested in confusing the comments from community $c_j$ are to the
LMCC. For this purpose, we consider the perplexity of the LMCC for a
comment coming from community $j$, defined as the exponential of the
entropy of the distribution \(D_j\), which is the (discrete)
probability distribution of predicted community for a random comment from
$c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
\[Ppl_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

@@

We see a positive correlation: this suggests that the more stable a
community is, the more difficult it is to identify comment as coming
from it. What this suggests is that stable communities tend to use a
varied and standard, subset of English, while unstable communities
resort to more formulaic language. This may suggests that to 'fit' in an
unstable community, one must use more obvious linguistic cues than in
a stable community. In fact, it is reasonable to think that a new
member will ostensibly use community-specifc language, but once
established in a community, a member would tend not to do so.
